Crunching Large Graphs with Commodity Processors

### 2011-01-26 doc/hotpar2011/softxmt-hotpar2011.tex
subsubroutine error
micro chages related work

%\paragraph{Single-node implementation.}  
In our single-node implementation, we implemented coroutines
as described, with a round-robin scheduler. Coroutines do local work,
issue a remote prefetch, and switch to the next in turn.  When the
coroutine is reactivated, it assumes the requested data is available
and issues a read.  If the data is not yet received and in cache, the core
(and all its coroutines) can stall.

%\paragraph{Multi-node ideas.}
As we develop our multi-node implementation, we will explore how the memory manager and coroutine scheduler may
interact, so that threads waiting on long-latency memory operations or
synchronization operations will be scheduled soon after their
operations have completed.

@@

%\paragraph{Single-node system.}
In our single-node implementation, we prefetch and yield before performing
a synchronization operation. We support only the synchronization
operations supported by our development platform; we have not
implemented full-bits yet.

%\paragraph{Multi-node ideas.}
In a multi-node implementation, synchronization operations can be delegated to a memory manager
thread or accelerator. Sychronization on remote data would be
performed by the remote memory manager. This may simplify the
implementation: if only a single core (or single accelerator) is
accessing the data being synchronized, the use of memory fences may be
reduced or eliminated.

@@

Our goal is to evaluate the feasibility of our proposed runtime. We
want to determine two things: whether coroutines can generate memory
concurrency while incurring minimal performance overhead, and whether the system can support the level of concurrency required to tolerate the latency that will be present in a multi-node system.

@@

Our evaulation consists of two parts. First, we demonstrate that the
runtime system can achieve the same performance as when 
explicit memory concurrency is available. Second, we look at the runtime
within the multi-node picture and show that it can tolerate large
latencies. In each part we begin by studying relevant aspects of our
test machine's memory system so we can interpret our runtime results.

\subsection{Single-node performance}

@@

\subsection{Base memory system performance}
\label{subsubsec:evalsinglebase}

@@

Figure~\ref{fig:multi-green} shows the result. For this
experiment, we limited memory concurrency to one concurrent miss per
coroutine, so the only source of memory concurrency is the use of
coroutines. We are able to obtain a rate of 275 \mrps\ with 48 concurrent misses, or 8 coroutines per core. These
results are very close to the ILP-only experiment, but more concurrent requests are required--about 48 versus 36. When the number of references per coroutine is larger than 1, fewer coroutines are required to reach full performance.

@@

% cache pressure figure
\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.42\textwidth]{figures/cache-pressure-edited.pdf}
  \end{center}
  \vspace{-0.3in}
  \caption{Cache pressure with coroutines, on six cores.}
  \label{fig:cache-pressure}
\end{figure}

@@

Figure~\ref{fig:cache-pressure} shows the result. As we would expect,
when there are no references to the local structures, the reference rates grow with the number of coroutines just as in
Figure~\ref{fig:multi-green}. For 32 local updates per traversal, while performance
decreases for large working set sizes and large numbers of coroutines,
we see speedup as the number of coroutines increases up to 128KB working set size with 8 coroutines per
core. \todo{preempt questions about drop at 16 which is much less than the about 50 needed per core by saying that the drop is likely the same drop effect seen earlier, which is likely to be prefetch squashing since it happens at number of line fill buffers}

@@

The results of the experiment are shown in
Figure~\ref{fig:network-delay}. We are still able to reach a
near-maximum rate of 273 \mrps\footnote{The throughput keeps increasing up to around 850 total concurrent references offered, far more than the 36 we know the chip to support. This is because we have modeled the network latency by forcing coroutines to wait the extra time before using the data, so the physical miss buffers on our machine are not tied up any longer than usual.}, but at least 2
concurrent references per coroutine are required. Running with one
reference per coroutine achieves 236 \mrps.

These results suggest that the runtime should be able to tolerate
large latency well enough to exceed the max issue rates of both modern
network interfaces (about 100 \mrps) \cite{mellanox:press} and the QPI
link used to reach the network interface.

@@

There has also been much work on user level threading, including the First-Class
User Level Threading project \cite{ult} and Capriccio \cite{capriccio}. We have a
different goal; we want many lightweight contexts to tolerate memory
latency.

Our goal of presenting a global view of distributed memory to the
programmer is shared by the PGAS community, and is used in languages
like Chapel \cite{chapel}, X10 \cite{X10}, and UPC \cite{upc}. They
obtain performance by minimizing references to remote nodes, whereas we
optimize for this case.

@@

We presented our plans for building a system for large graph
computation using commodity parts. We leverage decades-old ideas on
using a large amount of concurrency to tolerate latencies; but we do so
mostly in software. We developed a runtime system for latency
tolerance and our results showed that it is able to manage enough
concurrency to tolerate latencies long enough to amortize the cost of
expensive accesses to our large global memory system (high-latency,
but high-bandwidth).

>>>>

\paragraph{Single-node implementation.}  We implemented coroutines
with a round-robin scheduler. Coroutines do local work,
issue a remote prefetch, and yield. When a
coroutine is reactivated, it assumes the requested data is available
and issues a read.  If the data is not yet received and in cache, the core
(and all its coroutines) can stall.

\paragraph{Multi-node ideas.}
We will explore how the memory manager and coroutine scheduler may
interact, so that threads waiting on long-latency memory operations or
synchronization operations will be scheduled soon after their
operations have completed (and not before).

@@

\paragraph{Single-node system.}
Just as with reads and writes, we prefetch and yield before performing
a synchronization operation. We support only the synchronization
operations supported by our development platform; we have not
implemented full-bits yet.

\paragraph{Multi-node ideas.}
Synchronization operations can also be delegated to a memory manager
thread or accelerator. Sychronization on remote data would be
performed by the remote memory manager. This may simplify the
implementation: if only a single core (or single accelerator) is
accessing the data being synchronized, the use of memory fences may be
reduced or eliminated.

@@

Our goal is to evaluate the feasibility of our proposed runtime. We want to determine two things: whether coroutines can generate memory concurrency without incurring minimal performance overhead, and whether the system can support the level of concurrency required to tolerate the latency that will be present in a multi-node system.

@@

Our evaulation consists of two parts. First, we demonstrate that the
runtime system can achieve the same performance as when there is
explicit memory concurrency available. Second, we look at the runtime
within the multi-node picture and show that it can tolerate large
latencies. In each part we begin by studying relevant aspects of our
test machine's memory system so we can interpret our runtime results.

\subsection{Single-node}

@@

\subsubsection{Base memory system performance}
\label{subsubsec:evalsinglebase}

@@

Figure~\ref{fig:multi-green} shows the result. For this
experiment, we limited memory concurrency to one concurrent miss per
coroutine, so the only source of memory concurrency is the use of
coroutines. We are able to obtain a rate of 275 \mrps with 48 concurrent misses, or 8 coroutines per core. These
results are very close to the ILP-only experiment, but more concurrent requests are required--about 48 versus 36. When the number of references per coroutine is larger than 1, fewer coroutines are required to reach full performance.

@@

% cache pressure figure
\begin{figure}[h]
  \begin{center}
    \includegraphics[width=0.42\textwidth]{figures/cache-pressure-edited.pdf}
  \end{center}
  \caption{Cache pressure with coroutines, on six cores.}
  \label{fig:cache-pressure}
\end{figure}

@@

Figure~\ref{fig:cache-pressure} shows the result. As we would expect,
when there are no references to the local structures, the reference rates grow with the number of coroutines just as in
Figure~\ref{fig:multi-green}. For 32 local updates per traversal, while performance
decreases for large working set sizes and large numbers of coroutines,
we see speedup as number coroutines increases for up to 128KB working set size with 8 coroutines per
core. \todo{preempt questions about drop at 16 which is much less than the about 50 needed per core by saying that the drop is likely the same drop effect seen earlier, which is likely to be prefetch squashing since it happens at number of line fill buffers}

@@

The results of the experiment are shown in
Figure~\ref{fig:network-delay}. We are still able to reach a
near-maximum rate of 273 \mrps\footnote{The throughput keeps increasing up to around 850 total concurrent references offered, far more than the 36 we know the chip to support. This is because we have modeled the network latency by forcing coroutines to wait the extra time before using the data, so the physical miss buffers on our machine are not tied up any longer than usual.}, but at least 2
concurrent references per coroutine are required. Running with a single
reference per coroutine achieves 236 \mrps.

These results suggest that the runtime should be able to tolerate large latency well enough to exceed the max issue rates of both modern network interfaces (about 100 \mrps) \cite{mellanox:press} and the QPI link that connects the chip to the network interface.

@@

There has also been much work on user level threading; the First-Class
User Level Threading project \cite{ult} is one example. We have a
different goal; we want many lightweight contexts to tolerate memory
latency.

Our goal of presenting a global view of distributed memory to the
programmer is shared by the PGAS community, and is used in languages
like Chapel \cite{chapel}, X10 \cite{X10}, and UPC \cite{upc}. They
obtain performance by minimizing references to remote nodes, whereas we
design for remote references.

@@

We presented our plans on building a system for large graph
computation using commodity parts. We leverage decades-old ideas on
using a large ammount of concurrency to tolerate latencies; but do so
mostly in software. We developed a runtime system for latency
tolerance and our results showed that it is able to manage enough
concurrency to tolerate latencies long enough to amortize the cost of
expensive accesses to our large global memory system (high-latency,
but high-bandwidth).
